### 2018-12-09 paper/paper.tex
paper: Add initial draft and outline of paper

\documentclass[a4paper]{article}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}

\title{Neural Network Language Models for Automatic Speech Recognition}

\author{Aditya Kaushik, Eduardo Rosado, Thomas Spilsbury}

\date{\today}

\begin{document}
\maketitle

\begin{abstract}
In this paper we compare Recurrent Neural Network (RNN)
approximations of language models to traditional n-gram based models.
We evaluate different recurrent model architectures, hyperparameter configurations,
encoder and decode configurations and regularization and intialization procedures
and compare their performance on various performance and natural language
generation metrics. We find that in general, Long Short-Term Memory models
with embedding layers at the encoder stage outperforms other models
in terms of inference metrics and computational performance.
\end{abstract}

\section{Introduction}
\label{sec:introduction}

One important problem faced by Automatic Speech Recognition (ASR)
systems is transcribing utterances by speakers into intelligible written language.
In general, this is a difficult problem to solve due to the varying length
of utterances and ambiguity of classification between one utterance for another. The result
is that the recognized transcript is likely
to contain many words that are not
intelligible, even if the transcribed words
are the system's most likely direct
transcription from sounds to syllables.

As an alternative, we can make use of prior
information about the language to select
words from a vocabulary given some
observations of sounds or characters and the words
that have been predicted beforehand. The language model,
then, uses information about the sequence of prior
predicted words to give a probability distribution
for the next word, such that the ASR system would
pick the next word based both on what are the most
likely following words given a sequence as well
as the utterance that is made by the speaker.

Prior to the introduction of RNNs, most language models used n-grams
in Hidden Markov Model chains, where it was assumed that the $k$th
word depended only on the prior $n$ words. One key problem
with this model is that the context window is of a fixed size, leading
to a horizon problem. The Horizon problem occur because a if there were a word
in the word sequence which is extremely predictive of the $(n + 1)$th
word away from the current word, that would would simply not be taken
into account in that model. Worse still, the fixed size context window
means that unnecessary weight is put on words that happen to be in the context
window that may actually not be all that predictive were the context window
infinite.

In contrast, RNN's seek to solve this problem by introducing a Neural Network
architecture with recurrent connections, meaning that as the RNN makes predictions
over a sequence observations, it takes into account a hidden layer produced
by the previous observation (which in turn was a product of all the observations
before that). In effect, the RNN architecture encodes the contextual history
into the hidden layer vector allowing for an encoded approximation of infinite
contextual history.

Since the publication of \cite{Milkolov10}, new RNN approaches have shown
improvement on the state-of-the-art. We examine these new architectures and
techniques for future work discussed in the Milkolov paper and analyze whether their
results indicate that improvements can be made on the baseline set by the paper.

\section{Data}
\label{sec:data}

The first dataset we use is the WikiText dataset. This is a smaller dataset than
Gigaword and WSJ'92, but allowed us to
iterate faster and run more experiments.

This dataset is based on articles from
Wikipedia, containing 2,088,628 tokens
in the training set, 245,569 tokens in
the test set and 217,646 tokens in the
validation set.

Samples from the Dataset include:

"As with previous <unk> Chronicles games , Valkyria Chronicles III is a …"

"At Nintendo CEO Hiroshi Yamauchi 's request , Game Boy creator Gunpei Yokoi 's Nintendo R & D1 developed …"

"It was larger than the Scientific , at 73 by 155 by 34 millimetres ..."

This dataset contained quite a lot of
technical or domain specific
terminology due to the fact that the
sentences within were derived from
encyclopedic articles. This inherently
makes accurate prediction difficult, since
there is a higher likelihood that the
following word might be infrequent in the
corpus.

\section{Metrics}
\label{sec:metrics}

Models are evaluated by taking a sequence
of words from the validation set, removing
$k$ words from the end of the sentence and predicting
the next $k$ words from the existing $n - k$ word sequence.

\subsection{Perplexity}
\label{sec:perplexity}

Perplexity is a measure of "how generalized" the language
model is. A higher perplexity indicates a greater uncertainty
was encountered when predicting sequences of words. Formally,
it is the inverse probability of a text sequence normalized
by the number of tokens.

$$ PP(W) = \sqrt{\prod \frac{1}{P(w|w_1, ..., w_{i - 1}}} $$

Note that perplexity is an "intrinsic metric", it does not
have anything to do with the quality of the predicted
sentences, but instead is based on the certainty of the
model itself.

\subsection{ROUGE}
\label{sec:rouge}

In contrast, metrics such as ROUGE (which is an improvement
on BLEU) measure the quality of a sentence in terms of
their similarity to human generated language. These metrics
are called "extrinsic metrics".

The most basic form of such extrinsic metric is "word error
rate" (WER), defined as:

$$ WER = \frac{S + D + I}{N} $$, where $N$ is the number of
words and $S, D, I$ are substituions, deletions, insertions.

However, such a metric doesn't take int account the fact that
a recognized sequence may differ in length to a reference
sequence.

Improving on this situation, ROUGE measures based on the
longest matching subsequences, overlapping pairs and n-gram
co-occurrences.

\subsection{Naive Accuracy}
Accuracy is measured naively by comparing
a validation set sentence to a predicted sentence and
seeing how many words the model predicted correctly.

\section{Models}
\label{sec:models}

\subsection{RNN}
\label{sec:lstm}

A simple RNN (Elman network) as described in the Milkolov
paper consists of a single linear layer and non-linear
activation (in our case, the $tanh$ function, $ \frac{e^{2x} - 1}{e^{2x} + 1} $, with domain $[ -\infty, \infty ]$ and
range $ [-1, 1] $. The inputs to the network are
a hidden state vector, $h_{t}$, where the size is a hyperparameter
and the input encoding vector $i_{t}$. The output of the network
is the updated hidden state vector $h_{t + 1}$, which will will
be passed as the hidden state vector when processing the $i_{t + 1}$th
word, and the output encoding vector $o_{t}$, which can be
viewed as a discrete probability distribution for the predicted word
at $t$ by applying the softmax function $ \frac{e^x}{\sum e^x_i} $.

The trainable parameters of the RNN are the weights of the linear
layer which determine how information from the previous state of the
hidden state vector and the incoming word vector encoding are encoded
into both the updated hidden state vector and output word vector encoding.

The RNN is trained via "back-propagation through time" (BPTT), which is
effectively the same as the back-propgation algorithm proposed for feed-forward
neural networks, but applied to the recurrent structure of the RNN. In effect,
BPTT involves the same application of the chain rule to recurrent function
applications, and hidden layer states, but usually with some sort of threshold
to prevent computational complexity scaling with the number of tokens the
network has seen so far. In practice this means that the hidden layer encodes
a context window \emph{up to} to a certain length, but no longer than
that length.

\subsection{LSTM}
\label{sec:lstm}

\subsection{GRU}
\label{sec:gru}

\section{Encoding and Decoding}
\label{sec:encdec}

\subsection{Embeddings}
\label{sec:embedddings}

\subsection{Shortlists}
\label{sec:shortlists}

\subsection{Tied Weights}
\label{sec:tiedweights}

\section{Regularization}
\label{sec:regularization}

\subsection{Dropout}
\label{sec:dropout}

\section{Initialization}
\label{sec:initialization}

\subsection{Xavier}
\label{sec:xavier}

\section{Model Configurations}
\label{sec:configuration}

\section{Experimental Results}
\label{sec:results}

\section{Analysis}
\label{sec:analysis}

\section{Future Work}
\label{sec:future}

\begin{thebibliography}{9}
\bibitem{nano3}
  K. Grove-Rasmussen og Jesper Nygård,
  \emph{Kvantefænomener i Nanosystemer}.
  Niels Bohr Institute \& Nano-Science Center, Københavns Universitet

\end{thebibliography}
\end{document}

### 2018-12-10 paper/paper.tex
added some sections to the latex, made some small changes to current sections

@@
 \usepackage[colorinlistoftodos]{todonotes}
+\usepackage{graphicx}
+\graphicspath{ {./images/} }
 
@@
 We evaluate different recurrent model architectures, hyperparameter configurations,
-encoder and decode configurations and regularization and intialization procedures
-and compare their performance on various performance and natural language
+encoder and decoder configurations and regularization and intialization procedures
+while comparing their performance on various performance and natural language
 generation metrics. We find that in general, Long Short-Term Memory models
@@
 
-Since the publication of \cite{Milkolov10}, new RNN approaches have shown
+One of the first papers applying RNNs to Language Modeling was \cite{Milkolov10}. But since 2010, new RNN approaches have shown
 improvement on the state-of-the-art. We examine these new architectures and
@@
 $$ WER = \frac{S + D + I}{N} $$, where $N$ is the number of
-words and $S, D, I$ are substituions, deletions, insertions.
+words and $S, D, I$ are substitutions, deletions, insertions.
 
@@
 co-occurrences.
+
+\subsection{METEOR}
+\label{sec:meteor}
+Another metric trying to improve BLEU is METEOR.
+The novel thing about this metric is the use of Recall as well as Precision to compute the score.
+It computes the harmonic mean of Recall and Precision over uni-grams, the weight for recall being higher than Precision, meaning that false negatives (having a low probability for the true word trying to predict) will have a big impact on the metric (the score will be lower as the amount of false negatives increase).
+
+More formally, Precision over an uni-gram is computed as:
+$$P={\frac  {m}{w_{{t}}}}$$
+Where m is the number of uni-grams in the candidate translation that are also found in the reference translation, and $w_t$ is the number of uni-grams in the candidate translation.
+
+Recall is computed as:
+$$R={\frac  {m}{w_{{r}}}}$$
+Where m is as above, and $w_{{r}}$ is the number of uni-grams in the reference translation.
+Precision and recall are combined using the harmonic mean in the following fashion, with recall weighted 9 times more than precision:
+$$F_{{mean}}={\frac  {10PR}{R+9P}}$$
 
@@
 The RNN is trained via "back-propagation through time" (BPTT), which is
-effectively the same as the back-propgation algorithm proposed for feed-forward
+effectively the same as the back-propagation algorithm proposed for feed-forward
 neural networks, but applied to the recurrent structure of the RNN. In effect,
@@
 \label{sec:lstm}
+The LSTM cell builds upon the basic RNN cell and adds some gates in order to control how information is stored or deleted.
+These gates are element-wise operations, meaning that every value of the hidden state have a dedicated gate deciding the fate of that value in the next time step.
+All the gates are computed taking on account the last hidden state and the new input and they are applied onto the last hidden state.
 
+The first gate applied is a "forget" gate that regulates how much of the last hidden state is conserved in this time step. 
+The second one is a "storing" gate that decides how much of the new input is encoded into the hidden state.
+The last gate is an "output" gate which decides which parts of the hidden state are taken on account to compute the output.
+This output will pass through a softmax layer to get the prediction for that time step.
+
+Thanks to this procedure, hidden units can be ignored by the output gate and they can propagate forward in time without being multiplied by a weight matrix and applied an activation function at every time step, helping overcome the gradient vanishing problem.
+
+The forgetting and storing gates help encoding only relevant information into the hidden state, improving long-time dependencies recall.
 \subsection{GRU}
 \label{sec:gru}
+GRU is a simplified LSTM cell in which the forget and store gates are combined into one "update" gate with complementary gated values (they sum up to 1 for every hidden unit), this means that the hidden state will "forget" some information only if it is going to learn some new information.
 
@@
 \label{sec:embedddings}
+Traditionally, the Language Models were feeded each word as a One-Hot encoded vector as large as the vocabulary is.
+Since vocabularies can be really large (20k to 800k for languages with many compounded words such as Finnish), this presents a computational problem, large matrices multiplication is expensive.
+Word Embeddings provide an alternative procedure in which the words to be feeded to the network are encoded in an $n$ dimensional space.
+The embedding layer can learn a word encoding such that words related to each other semanticaly, get similar encoded vectors after the transformation.
+Even some basic arithmetic is possible. For example, if you subtract the vector for "queen" from the vector for "king" you get the vector for "man".
+
+\includegraphics[width=\textwidth]{01}
 

### 2018-12-10 paper/paper.tex
paper: Reflow contents of LSTM/encodings/etc

@@
 \label{sec:meteor}
-Another metric trying to improve BLEU is METEOR.
-The novel thing about this metric is the use of Recall as well as Precision to compute the score.
-It computes the harmonic mean of Recall and Precision over uni-grams, the weight for recall being higher than Precision, meaning that false negatives (having a low probability for the true word trying to predict) will have a big impact on the metric (the score will be lower as the amount of false negatives increase).
+Another metric trying to improve BLEU is
+METEOR. The novel thing about this metric is the use of Recall as well as
+Precision to compute the score. It computes the harmonic mean of Recall and
+Precision over uni-grams, the weight for recall being higher than Precision,
+meaning that false negatives (having a low probability for the true word trying
+to predict) will have a big impact on the metric (the score will be lower as the
+amount of false negatives increase).
 
-More formally, Precision over an uni-gram is computed as:
-$$P={\frac  {m}{w_{{t}}}}$$
-Where m is the number of uni-grams in the candidate translation that are also found in the reference translation, and $w_t$ is the number of uni-grams in the candidate translation.
+More formally, Precision over an uni-gram is computed as: $$P={\frac
+{m}{w_{{t}}}}$$ Where m is the number of uni-grams in the candidate translation
+that are also found in the reference translation, and $w_t$ is the number of
+uni-grams in the candidate translation.
 
-Recall is computed as:
-$$R={\frac  {m}{w_{{r}}}}$$
-Where m is as above, and $w_{{r}}$ is the number of uni-grams in the reference translation.
-Precision and recall are combined using the harmonic mean in the following fashion, with recall weighted 9 times more than precision:
-$$F_{{mean}}={\frac  {10PR}{R+9P}}$$
+Recall is computed as: $$R={\frac {m}{w_{{r}}}}$$ Where m is as above, and
+$w_{{r}}$ is the number of uni-grams in the reference translation. Precision and
+recall are combined using the harmonic mean in the following fashion, with
+recall weighted 9 times more than precision: $$F_{{mean}}={\frac {10PR}{R+9P}}$$
 
@@
 
-\subsection{LSTM}
-\label{sec:lstm}
-The LSTM cell builds upon the basic RNN cell and adds some gates in order to control how information is stored or deleted.
-These gates are element-wise operations, meaning that every value of the hidden state have a dedicated gate deciding the fate of that value in the next time step.
-All the gates are computed taking on account the last hidden state and the new input and they are applied onto the last hidden state.
+\subsection{LSTM} \label{sec:lstm}
+The LSTM cell builds upon the basic RNN cell
+and adds some gates in order to control how information is stored or deleted.
+These gates are element-wise operations, meaning that every value of the hidden
+state have a dedicated gate deciding the fate of that value in the next time
+step. All the gates are computed taking on account the last hidden state and the
+new input and they are applied onto the last hidden state.
 
-The first gate applied is a "forget" gate that regulates how much of the last hidden state is conserved in this time step. 
-The second one is a "storing" gate that decides how much of the new input is encoded into the hidden state.
-The last gate is an "output" gate which decides which parts of the hidden state are taken on account to compute the output.
-This output will pass through a softmax layer to get the prediction for that time step.
+The first gate applied is a "forget" gate that regulates how much of the last
+hidden state is conserved in this time step. The second one is a "storing" gate
+that decides how much of the new input is encoded into the hidden state. The
+last gate is an "output" gate which decides which parts of the hidden state are
+taken on account to compute the output. This output will pass through a softmax
+layer to get the prediction for that time step.
 
-Thanks to this procedure, hidden units can be ignored by the output gate and they can propagate forward in time without being multiplied by a weight matrix and applied an activation function at every time step, helping overcome the gradient vanishing problem.
+Thanks to this procedure, hidden units can be ignored by the output gate and
+they can propagate forward in time without being multiplied by a weight matrix
+and applied an activation function at every time step, helping overcome the
+gradient vanishing problem.
 
-The forgetting and storing gates help encoding only relevant information into the hidden state, improving long-time dependencies recall.
-\subsection{GRU}
-\label{sec:gru}
-GRU is a simplified LSTM cell in which the forget and store gates are combined into one "update" gate with complementary gated values (they sum up to 1 for every hidden unit), this means that the hidden state will "forget" some information only if it is going to learn some new information.
+The forgetting and storing gates help encoding only relevant information into
+the hidden state, improving long-time dependencies recall. \subsection{GRU}
+\label{sec:gru} GRU is a simplified LSTM cell in which the forget and store
+gates are combined into one "update" gate with complementary gated values (they
+sum up to 1 for every hidden unit), this means that the hidden state will
+"forget" some information only if it is going to learn some new information.
 
@@
 
-\subsection{Embeddings}
-\label{sec:embedddings}
-Traditionally, the Language Models were feeded each word as a One-Hot encoded vector as large as the vocabulary is.
-Since vocabularies can be really large (20k to 800k for languages with many compounded words such as Finnish), this presents a computational problem, large matrices multiplication is expensive.
-Word Embeddings provide an alternative procedure in which the words to be feeded to the network are encoded in an $n$ dimensional space.
-The embedding layer can learn a word encoding such that words related to each other semanticaly, get similar encoded vectors after the transformation.
-Even some basic arithmetic is possible. For example, if you subtract the vector for "queen" from the vector for "king" you get the vector for "man".
+\subsection{Embeddings} \label{sec:embedddings} Traditionally, the Language
+Models were feeded each word as a One-Hot encoded vector as large as the
+vocabulary is. Since vocabularies can be really large (20k to 800k for languages
+with many compounded words such as Finnish), this presents a computational
+problem, large matrices multiplication is expensive. Word Embeddings provide an
+alternative procedure in which the words to be feeded to the network are encoded
+in an $n$ dimensional space. The embedding layer can learn a word encoding such
+that words related to each other semanticaly, get similar encoded vectors after
+the transformation. Even some basic arithmetic is possible. For example, if you
+subtract the vector for "queen" from the vector for "king" you get the vector
+for "man".
 

### 2018-12-10 paper/paper.tex
paper: Add additional information LSTM, GRU and embeddings

--- 2018-12-10 paper/paper.tex
@@
 
+Unfortunately, Elman Network RNNs are not able to retain all the optimal
+contextual in practice. The reasons for this are given by \cite{Bengio94},
+but in summary the issue is that as we backpropagate through time the gradients
+from states "further back" in the history start to get vanishingly small, meaning
+that they can be easily disrupted by noise or newer information. This means
+that the hidden layer doesn't retain long-term information since it is easily
+disrupted by new information.
+
 \subsection{LSTM}
 \label{sec:lstm}
-The LSTM cell builds upon the basic RNN cell and adds some gates in order to control how information is stored or deleted.
-These gates are element-wise operations, meaning that every value of the hidden state have a dedicated gate deciding the fate of that value in the next time step.
-All the gates are computed taking on account the last hidden state and the new input and they are applied onto the last hidden state.
+In contrast to Elman RNNs, LSTM cells remember long-term information "by default" \cite{hochreiter97}.
+The hidden state isn't regenerated from the old hidden state and new information coming
+through the input, but rather is updated by deleting information and adding new
+information.
 
-The first gate applied is a "forget" gate that regulates how much of the last hidden state is conserved in this time step. 
-The second one is a "storing" gate that decides how much of the new input is encoded into the hidden state.
-The last gate is an "output" gate which decides which parts of the hidden state are taken on account to compute the output.
-This output will pass through a softmax layer to get the prediction for that time step.
+The LSTM cell uses four "gates" to manage the state of the memory. Each of these
+gates comprises a set of weights and activations, descibed below \cite{colah2015}.
 
-Thanks to this procedure, hidden units can be ignored by the output gate and they can propagate forward in time without being multiplied by a weight matrix and applied an activation function at every time step, helping overcome the gradient vanishing problem.
+The LSTM cell builds upon the basic RNN cell
+and adds some gates in order to control how information is stored or deleted.
+These gates are element-wise operations, meaning that every value of the hidden
+state have a dedicated gate deciding the fate of that value in the next time
+step. All the gates are computed taking on account the last hidden state and the
+new input and they are applied onto the last hidden state.
 
-The forgetting and storing gates help encoding only relevant information into the hidden state, improving long-time dependencies recall.
+The first gate applied is a "forget" gate that regulates how much of the last
+hidden state is conserved in this time step. Say that the hidden state contained some
+sort of representation as to a place or a point in time, but a new place or point in time arrived.
+The forget gate would, in theory, be optimized to notice the new place or point in time and delete
+the old place or point in time, otherwise leaving the contextual information unchanged. The output
+of the forget gate is activated by the sigmoid function which creates a mask of what information
+to retain from the hidden state ($f_t$).
+
+The second one is an "input" gate that decides how much of the new input is encoded into the hidden state.
+The activations for the gate by sigmoid essentially form a "mask" which the new input feature vector
+$C_t$ gets multiplied by (in order to select the relevant features) ($i_t$).
+
+In order to update the hidden state, we multiply the old hidden state ($h_{t - 1}$) by $f_t$
+(masking out the information we want to forget) and multiply the input state by $i_t$, masking in
+the information we want to keep, and add them together.
+
+The last gate is an "output" gate which decides which parts of the hidden state are
+taken on account to compute the output. The hidden state is first activated by sigmoid to determin
+the mask ($o_t$), then the hidden layer is activated by tanh to push the values between $[-1, 1]$, then
+finally multipled by the mask $o_t$ in order to determine the output of the network.
+
+The end result with such a variation is that the hidden state is "protected" from perturbation
+by virtue of being combined through a single linear layer and activation layer with the input
+state on every timestep. Instead, the hidden state is more "explicitly" modified when the
+optimized input and forget gates explicitly indicate that information should be added or removed
+to the state \cite{hochreiter97}.
+
+The output from the output gate passes through a softmax layer to get the prediction for that time step.
+
+Thanks to this procedure, hidden units can be ignored by the output gate and
+they can propagate forward in time without being multiplied by a weight matrix
+and applied an activation function at every time step, helping overcome the
+gradient vanishing problem.
+
+The forgetting and storing gates help encoding only relevant information into
+the hidden state, improving long-time dependencies recall.
+
+Note that the hidden state accuracy benefits here come with a computational compleixty cost - every
+LSTM cell has three internal linear layers and five activations, which we pay for both in the forward
+pass and in backpropagation through time.
+
 \subsection{GRU}
 \label{sec:gru}
-GRU is a simplified LSTM cell in which the forget and store gates are combined into one "update" gate with complementary gated values (they sum up to 1 for every hidden unit), this means that the hidden state will "forget" some information only if it is going to learn some new information.
+
+In contrast with the LSTM model, Gated Recurrent Unit (GRU) \cite{cho2014} is a simplified
+LSTM cell in which the forget and store gates are combined into one "update" gate with complementary
+gated values (they sum up to 1 for every hidden unit). This means that the hidden state will
+"forget" some information only if it is going to learn some new information. The GRU also
+introduces a "reset gate". The "reset gate" is computed by taking the sigmoid of the sum of weighted
+inputs and hidden state $\sigma ([\bold W_r \bold x]_j + [\bold U_r \bold_{h - 1}])$.
+
+When the "reset gate" is close to zero, the entire hidden state is thrown away and resetted with
+the current input gate state.
+
+The net effect is that the GRU can learn short term dependencies between inputs
+over different timescales. The units that capture short term dependencies frequently
+activate the reset gate, whereas the units that learn long term dependencies have
+more active update gates \cite{cho2014}.
 
@@
 \label{sec:embedddings}
-Traditionally, the Language Models were feeded each word as a One-Hot encoded vector as large as the vocabulary is.
-Since vocabularies can be really large (20k to 800k for languages with many compounded words such as Finnish), this presents a computational problem, large matrices multiplication is expensive.
-Word Embeddings provide an alternative procedure in which the words to be feeded to the network are encoded in an $n$ dimensional space.
-The embedding layer can learn a word encoding such that words related to each other semanticaly, get similar encoded vectors after the transformation.
-Even some basic arithmetic is possible. For example, if you subtract the vector for "queen" from the vector for "king" you get the vector for "man".
+
+Traditionally, the Language Models were feeded each word as a One-Hot encoded vector as large as the
+vocabulary is. Since vocabularies can be really large (20k to 800k for languages
+with many compounded words such as Finnish), this presents a computational
+problem, large matrices multiplication is expensive. Word Embeddings provide an
+alternative procedure in which the words to be feeded to the network are encoded
+in an $n$ dimensional space. The embedding layer can learn a word encoding such
+that words related to each other semanticaly, get similar encoded vectors after
+the transformation. Even some basic arithmetic is possible. For example, if you
+subtract the vector for "queen" from the vector for "king" you get the vector
+for "man".
+
+Word embeddings can be learned end-to-end with a single training
+process, as long as there is some task giving the words a representation in a domain,
+for instance, predicting the next word, which a language model is designed to do. This
+means that the word embeddings need not be trained separately but can be learned
+as part of the process of training the entire language model as a separate layer.
+
+Word embeddings also have the distinct advantage that they can reduce both training
+and inference complexity because fewer multiplications need to be done at
+the RNN, LSTM or GRU cell phase both at the inference and training time. The
+comptutation of the embedding layer can be heavily optimized because the relevant
+input will only have a non-zero value in a single vector component.
 
@@
 
+\bibitem{Bengio94}
+  http://ai.dinfo.unifi.it/paolo//ps/tnn-94-gradient.pdf
+
+\bibitem{colah2015}
+  http://colah.github.io/posts/2015-08-Understanding-LSTMs/
+
+\bibitem{hochreiter97}
+  http://www.bioinf.jku.at/publications/older/2604.pdf
+
+\bibitem{cho2014}
+  https://arxiv.org/pdf/1406.1078v3.pdf
 \end{thebibliography}

### 2018-12-10 paper/paper.tex
paper: Remove extra whitespace
paper: Add shortlists section

@@
 
+Like embeddings, shortlists also provide a way to reduce the computational
+complexity, but this time at the output layer instead of the inference layer
+\cite{schwenk05}. In the shortlist we pick the $n$ most frequent words in the
+corpus and $n$ output units to those $n$ words in the corpus, predicting every
+other uncommon word as an unknown token. However, input encodings still
+allow for representations of those words deemed uncommon, either by
+embeddings or one-hot encoding.
+
+The overall theory behind shortlists is that any optimal model operates
+on a pareto principle, 20% of the words which are the most frequent will
+be predicted roughly 80% of the time. Since the computation of the output
+encodings of height $|V|$ is a fixed cost on every training and inference
+cycle, the user of the model is paying for the capacity to predict words which
+will seldomly be predicted. A good shortlist should be able to substantially
+reduce the computational complexity of the model whilst still retaining a
+relatively high degree of accuracy.
+
 \subsection{Tied Weights}

### 2018-12-10 paper/paper.tex
paper: Add section on regularization and dropout

@@
 
+The Milkolov paper flags that performance on the validation could not be
+improved by using "standard regularization procedures" (we assume that
+the paper must have been referring to L1 or L2 regualrization based on the
+techniques in existence at the time) \cite{Milkolov10}. We propose that a
+popular regularization technique known as "Dropout" may be of some assitance.
+
 \subsection{Dropout}
 \label{sec:dropout}
+
+Dropout \cite{hinton2012} is the probabilistic approximation of an ensemble
+of the power-set of weights in a Neural Network. In the words of Hinton, it
+reduces overfitting by reducing "complex co-adaptations on the training data",
+meaning that the network does not depend too much on the activation of
+certain sets of units in order to produce the majority of activations, instead
+ensuring that every unit is utilized roughly equally. In the case of language models,
+this should in principle prevent the network from just recalling the training data.
 

### 2018-12-10 paper/paper.tex
paper: Citations for Milkolov, Schwenk, etc
paper: Merge results and configuration

@@
 
-\section{Model Configurations}
-\label{sec:configuration}
+\section{Model Configurations and Results}
+\label{sec:results}
 
-\section{Experimental Results}
-\label{sec:results}
+We ran two sets of experiments on each dataset. Because of the difference in sizes
+between the Gigaword and WikiText datasets we were able to run more tests on the
+WikiText dataset and scale a few of those results up to Gigaword.
 
@@
 
+\bibitem{Milkolov10}
+  Milkolov T., et al, Recurrent neural network based language model, INTERSPEECH 2010
+
 \bibitem{Bengio94}
@@
   https://arxiv.org/pdf/1406.1078v3.pdf
+
+\bibitem{schwenk05}
+  ftp://tlp.limsi.fr/public/emnlp05.pdf. Accessed 6 Nov. 2018.
+
+\bibitem{hinton2012}
+  https://arxiv.org/pdf/1207.0580.pdf
+
 \end{thebibliography}
+
 \end{document}

### 2018-12-10 paper/paper.tex
paper: Add more detail on WikiText and Gigaword, including EDA images

@@
 
+\section{WikiText}
+\label{sec:wikitext}
+
 The first dataset we use is the WikiText dataset. This is a smaller dataset than
@@
 
-This dataset is based on articles from
-Wikipedia, containing 2,088,628 tokens
-in the training set, 245,569 tokens in
-the test set and 217,646 tokens in the
-validation set.
+This dataset is based on articles from Wikipedia, containing 1,756,924 tokens in
+the training set, 184,354 tokens in the validation set and 207,080 tokens
+in the test set. The combined dictionary for all three sets contains 28,755
+unique tokens.
+
+\begin{figure}[h]
+\includegraphics{sr-eda-wikitext-train-words}
+\includegraphics{sr-eda-wikitext-valid-words}
+\centering
+\caption{Training and Validation words in the WikiText dataset}
+\end{figure}
+
+Notice that because of the large amount of domain specific language that is used
+due to the fact that the sentences within were derived from encyclopedic
+articles, the dataset contains a large number of "unknown" tokens. An "unknown"
+token, represents any token with a frequency below the dataset threshold. This
+inherently makes accurate prediction difficult, since there is a higher
+likelihood that the following word might be infrequent in the corpus.
 
@@
 
-This dataset contained quite a lot of
-technical or domain specific
-terminology due to the fact that the
-sentences within were derived from
-encyclopedic articles. This inherently
-makes accurate prediction difficult, since
-there is a higher likelihood that the
-following word might be infrequent in the
-corpus.
+Also notice the most frequent bigrams and trigrams of the dataset
+
+\begin{figure}[h]
+\includegraphics{sr-eda-wikitext-bigrams}
+\includegraphics{sr-eda-wikitext-trigrams}
+\centering
+\caption{Training set Bigrams and Trigrams in the WikiText dataset}
+\end{figure}
+
+The bigrams and trigrams in the dataset indicate that the dataset is, as expected,
+most describing phenomena or objects, with large frequencies for bigrams
+such as "in the" or "of the". The trigrams also reveal that domain specific
+descriptions-in-general are frequently occurring, ("unk unk unk").
+
+\section{Gigaword}
+\label{sec:gigaword}
+
+The second dataset we used was Gigaword. This dataset is based on ____,
+containing 110,382,986 tokens in the training set, 549,0454 tokens in the
+validation set and ____ tokens in the test set. The combined dictionary
+contains ____ unique tokens.
+
+Samples from the Dataset include:
+
+____
+
+\begin{figure}[h]
+\includegraphics{sr-eda-gigaword-train-words}
+\includegraphics{sr-eda-gigaword-valid-words}
+\centering
+\caption{Training and Validation words in the Gigaword dataset}
+\end{figure}
+
+In contrast to WikiText, there are a smaller number of domain-specific
+words and instead more words similar to those typically used in news
+articles, "said", "government", "stock", "tuesday", etc.
+
+\begin{figure}[h]
+\includegraphics{sr-eda-gigaword-bigrams}
+\includegraphics{sr-eda-gigaword-trigrams}
+\centering
+\caption{Bigrams and Trigrams words in the Gigaword training set}
+\end{figure}
+
+Also notice that bigrams and trigrams for this dataset reveal that a
+large number of articles are talking about the same thing, for instance
+"the united states", "the european union" and "george w bush".
 

### 2018-12-10 paper/paper.tex
paper: Add performance metrics and comment for wikitext

@@
 
+\section{Performance by Model Architecture}
+\label{sec:perf_by_model_arch}
+
+\begin{figure}[h]
+\includegraphics{sr-perf-by-model-architecture}
+\centering
+\caption{Validation Set Performance by Model Architecture on WikiText}
+\end{figure}
+
+Comparing each of the three models with 200 hidden units, 2 layers and
+no dropout or embeddings, trained to five epochs in batches of 200 sentences at a
+time, we see that the LSTM model generally outperforms all three on all metrics.
+It has the lowest overall perplexity, highest METEOR score after five iterations
+and highest overall naive accuracy, scoring roughly one half a percentage point
+higher than the other two models.
+
+\section{Performance by number of hidden layers}
+\label{sec:perf_by_n_layers}
+
+\begin{figure}[h]
+\includegraphics{sr-perf-by-hidden-lstm}
+\includegraphics{sr-perf-by-hidden-gru}
+\includegraphics{sr-perf-by-hidden-rnn}
+\centering
+\caption{Validation Set Performance by number of Hidden Layers on WikiText}
+\end{figure}
+
+We measured differences in performance with both one and two hidden layers
+on the RNN cell for each model architecture with 200 hidden units per cell, trained
+to five epochs in batches of 200 sentences at a time.
+
+In general, we find that increasing the number of hidden layers on the network
+increases performance on all metrics, probably because the model is able to
+encode more information from the training set in its layers recurrent layers.
+
+We notice that in the Elman Network RNN training was generally unstable and
+increasing the number of layers gave inconclusive results.
+
+\section{Performance by number of hidden units}
+\label{sec:perf_by_n_hidden}
+
+\begin{figure}[h]
+\includegraphics{sr-perf-by-hidden-units-lstm}
+\includegraphics{sr-perf-by-hidden-units-gru}
+\includegraphics{sr-perf-by-hidden-units-rnn}
+\centering
+\caption{Validation Set Performance by number of Hidden Units on WikiText}
+\end{figure}
+
+We measured differences in performance with 60, 90, 150 and 200 hidden units per
+RNN cell for each model architecture with 1 RNN cell, trained
+to five epochs in batches of 200 sentences at a time.
+
+In general we found that like increasing the number of layers,
+increasing the number of hidden units per layer can help with overall accuracy,
+but not as much as increasing the number of layers does. In fact, having a fewer
+number of hidden units may act as a regularizer which would help to prevent
+overfitting and improve accuracy on the validation data.
+
+\section{Performance by Dropout}
+\label{sec:perf_by_dropout}
+
+\begin{figure}[h]
+\includegraphics{sr-perf-by-dropout-lstm}
+\includegraphics{sr-perf-by-dropout-gru}
+\includegraphics{sr-perf-by-dropout-rnn}
+\centering
+\caption{Validation Set Performance by Dropout Probability on WikiText}
+\end{figure}
+
+We measured differences in performance with dropout probabilities
+0.0, 0.01, 0.05, 0.1, 0.2 per RNN cell for each model architecture with
+1 RNN cell, 200 hidden units, trained to five epochs in batches of 200
+sentences at a time.
+
+For GRU cells, we find that small amounts of dropout (0.01-0.05) can act
+as an effective regularizer and improve accuracy and ROUGE scores on unseen
+data by a few tenths of a percentage point. For LSTM cells we find that
+dropout did not assist all that much, indicating that perhaps those models were
+not already starting to overfit by five epochs. For RNN cells, we observed
+that again, training was quite unstable that it was difficult to observe what
+difference, if any, dropout had on the metrics.
+
+\section{Performance by Embedding Dimensions}
+\label{sec:perf_by_embedding}
+
+\begin{figure}[h]
+\includegraphics{sr-perf-by-embedding-lstm}
+\includegraphics{sr-perf-by-embedding-gru}
+\includegraphics{sr-perf-by-embedding-rnn}
+\centering
+\caption{Validation Set Performance by Embedding Dimension on WikiText}
+\end{figure}
+
+We measured differences in performance with embedding sizes 0, 50,
+100, 200 and 500 at the encoder layer for each model architecture with
+1 RNN cell, 200 hidden units, trained to five epochs in batches of 200
+sentences at a time.
+
+On both LSTM and GRU models, any use of encoder embeddings noticably outperformed
+one-hot encoding on almost all metrics, except METEOR for LSTMs. It seems to be
+the case that the learned embeddings encoded useful information about the
+non-temporally dependent relationships between words. For Elman RNN cells, we
+again observed that because training was so unstable, embeddings did not
+appear to make much of a difference to accuracy, even if they result in
+much lower computational complexity.
+
+\section{Performance by Shortlist Dimension}
+\label{sec:perf_by_shortlist}
+
+\begin{figure}[h]
+\includegraphics{sr-perf-by-shortlist-lstm}
+\includegraphics{sr-perf-by-shortlist-gru}
+\includegraphics{sr-perf-by-shortlist-rnn}
+\centering
+\caption{Validation Set Performance by Shortlist Dimension on WikiText}
+\end{figure}
+
+We measured differences in performance with shortlist sizes 100, 1000, 5000, 1000
+at the decoder layer for each model architecture with 1 RNN cell, 200 hidden units and
+an input encoder embedding size of 500, trained to five epochs in batches of 200 sentences at a time.
+
+On WikiText, note that the original corpus contained 28,755 words in its
+dictionary and achieved accuracy scores of 0.921 for the LSTM, 0.918 for the GRU
+and 0.911 for Elman RNN.
+
+By contrast, shortlists for LSTM, GRU and Elman RNNs
+achieve accuracy scores between 0.910 and to 0.911,
+with larger shortlists marginally outperforming smaller
+shortlists. This indicates that a large proportion of
+the words being predicted on the non-shortlisted
+models are likely to be very frequent words, meaning
+that if we \emph{only} allow the network to predict
+frequent words by shortlisting then we can still get
+slightly worse but similar performance.
+
+\section{Tied Weights}
+\label{sec:perf_by_tiedweights}
+
+\begin{figure}[h]
+\includegraphics{sr-perf-by-model-tied-weights}
+\centering
+\caption{Validation Set Performance by Model on WikiText for Tied Weights}
+\end{figure}
+
+\section{Xavier Initialization}
+\label{sec:perf_by_xavier_init}
+
+\begin{figure}[h]
+\includegraphics{sr-perf-by-model-xavier-init}
+\centering
+\caption{Validation Set Performance by Model on WikiText for Xavier Initialization}
+\end{figure}
+
+\section{Kaiming Initialization}
+\label{sec:perf_by_kaiming_init}
+
+\begin{figure}[h]
+\includegraphics{sr-perf-by-model-kaiming-init}
+\centering
+\caption{Validation Set Performance by Model on WikiText for Xavier Initialization}
+\end{figure}
+
 \section{Analysis}

### 2018-12-10 paper/paper.tex
paper: Replace underscores with XXX

@@
 
-The second dataset we used was Gigaword. This dataset is based on ____,
+The second dataset we used was Gigaword. This dataset is based on XXX,
 containing 110,382,986 tokens in the training set, 549,0454 tokens in the
-validation set and ____ tokens in the test set. The combined dictionary
-contains ____ unique tokens.
+validation set and XXX tokens in the test set. The combined dictionary
+contains XXX unique tokens.
 
@@
 
-____
+XXX
 

### 2018-12-10 paper/paper.tex
paper: Resize all the images to be the correct size
They still don't flow correctly

@@
 
-\begin{figure}[h]
-\includegraphics{sr-eda-wikitext-train-words}
-\includegraphics{sr-eda-wikitext-valid-words}
+\begin{figure}[ht]
+\includegraphics[width=0.7\columnwidth]{sr-eda-wikitext-train-words}
+\includegraphics[width=0.7\columnwidth]{sr-eda-wikitext-valid-words}
 \centering
@@
 
-\begin{figure}[h]
-\includegraphics{sr-eda-wikitext-bigrams}
-\includegraphics{sr-eda-wikitext-trigrams}
+\begin{figure}[ht]
+\includegraphics[width=0.7\columnwidth]{sr-eda-wikitext-bigrams}
+\includegraphics[width=0.7\columnwidth]{sr-eda-wikitext-trigrams}
 \centering
@@
 
-\begin{figure}[h]
-\includegraphics{sr-eda-gigaword-train-words}
-\includegraphics{sr-eda-gigaword-valid-words}
+\begin{figure}[ht]
+\includegraphics[width=0.7\columnwidth]{sr-eda-gigaword-train-words}
+\includegraphics[width=0.7\columnwidth]{sr-eda-gigaword-valid-words}
 \centering
@@
 
-\begin{figure}[h]
-\includegraphics{sr-eda-gigaword-bigrams}
-\includegraphics{sr-eda-gigaword-trigrams}
+\begin{figure}[ht]
+\includegraphics[width=0.7\columnwidth]{sr-eda-gigaword-bigrams}
+\includegraphics[width=0.7\columnwidth]{sr-eda-gigaword-trigrams}
 \centering
@@
 
-\includegraphics[width=\textwidth]{01}
+\includegraphics[width=0.7\columnwidth][width=\textwidth]{01}
 
@@
 
-\begin{figure}[h]
-\includegraphics{sr-perf-by-model-architecture}
+\begin{figure}[ht]
+\includegraphics[width=0.7\columnwidth]{sr-perf-by-model-architecture}
 \centering
@@
 
-\begin{figure}[h]
-\includegraphics{sr-perf-by-hidden-lstm}
-\includegraphics{sr-perf-by-hidden-gru}
-\includegraphics{sr-perf-by-hidden-rnn}
+\begin{figure}[ht]
+\includegraphics[width=0.7\columnwidth]{sr-perf-by-hidden-lstm}
+\includegraphics[width=0.7\columnwidth]{sr-perf-by-hidden-gru}
+\includegraphics[width=0.7\columnwidth]{sr-perf-by-hidden-rnn}
 \centering
@@
 
-\begin{figure}[h]
-\includegraphics{sr-perf-by-hidden-units-lstm}
-\includegraphics{sr-perf-by-hidden-units-gru}
-\includegraphics{sr-perf-by-hidden-units-rnn}
+\begin{figure}[ht]
+\includegraphics[width=0.7\columnwidth]{sr-perf-by-hidden-units-lstm}
+\includegraphics[width=0.7\columnwidth]{sr-perf-by-hidden-units-gru}
+\includegraphics[width=0.7\columnwidth]{sr-perf-by-hidden-units-rnn}
 \centering
@@
 
-\begin{figure}[h]
-\includegraphics{sr-perf-by-dropout-lstm}
-\includegraphics{sr-perf-by-dropout-gru}
-\includegraphics{sr-perf-by-dropout-rnn}
+\begin{figure}[ht]
+\includegraphics[width=0.7\columnwidth]{sr-perf-by-dropout-lstm}
+\includegraphics[width=0.7\columnwidth]{sr-perf-by-dropout-gru}
+\includegraphics[width=0.7\columnwidth]{sr-perf-by-dropout-rnn}
 \centering
@@
 
-\begin{figure}[h]
-\includegraphics{sr-perf-by-embedding-lstm}
-\includegraphics{sr-perf-by-embedding-gru}
-\includegraphics{sr-perf-by-embedding-rnn}
+\begin{figure}[ht]
+\includegraphics[width=0.7\columnwidth]{sr-perf-by-embedding-lstm}
+\includegraphics[width=0.7\columnwidth]{sr-perf-by-embedding-gru}
+\includegraphics[width=0.7\columnwidth]{sr-perf-by-embedding-rnn}
 \centering
@@
 
-\begin{figure}[h]
-\includegraphics{sr-perf-by-shortlist-lstm}
-\includegraphics{sr-perf-by-shortlist-gru}
-\includegraphics{sr-perf-by-shortlist-rnn}
+\begin{figure}[ht]
+\includegraphics[width=0.7\columnwidth]{sr-perf-by-shortlist-lstm}
+\includegraphics[width=0.7\columnwidth]{sr-perf-by-shortlist-gru}
+\includegraphics[width=0.7\columnwidth]{sr-perf-by-shortlist-rnn}
 \centering
@@
 
-\begin{figure}[h]
-\includegraphics{sr-perf-by-model-tied-weights}
+\begin{figure}[ht]
+\includegraphics[width=0.7\columnwidth]{sr-perf-by-model-tied-weights}
 \centering
@@
 
-\begin{figure}[h]
-\includegraphics{sr-perf-by-model-xavier-init}
+\begin{figure}[ht]
+\includegraphics[width=0.7\columnwidth]{sr-perf-by-model-xavier-init}
 \centering
@@
 
-\begin{figure}[h]
-\includegraphics{sr-perf-by-model-kaiming-init}
+\begin{figure}[ht]
+\includegraphics[width=0.7\columnwidth]{sr-perf-by-model-kaiming-init}
 \centering

### 2018-12-11 paper/paper.tex
paper: Move things into subsections
added xavier init. and tied weights to the paper and slides

@@
 
-\begin{figure}[ht]
+\begin{figure}[!ht]
 \includegraphics[width=0.7\columnwidth]{sr-eda-wikitext-train-words}
@@
 
-"At Nintendo CEO Hiroshi Yamauchi 's request , Game Boy creator Gunpei Yokoi 's Nintendo R & D1 developed …"
+"At Nintendo CEO Hiroshi Yamauchi 's request , Game Boy creator Gunpei Yokoi 's Nintendo R \& D1 developed …"
 
@@
 
-\begin{figure}[ht]
+\begin{figure}[!ht]
 \includegraphics[width=0.7\columnwidth]{sr-eda-wikitext-bigrams}
@@
 
-\begin{figure}[ht]
+\begin{figure}[!ht]
 \includegraphics[width=0.7\columnwidth]{sr-eda-gigaword-train-words}
@@
 
-\begin{figure}[ht]
+\begin{figure}[!ht]
 \includegraphics[width=0.7\columnwidth]{sr-eda-gigaword-bigrams}
@@
 introduces a "reset gate". The "reset gate" is computed by taking the sigmoid of the sum of weighted
-inputs and hidden state $\sigma ([\bold W_r \bold x]_j + [\bold U_r \bold_{h - 1}])$.
+inputs and hidden state
+ $\sigma ([ W_r d x]_j + [ U_r  h - 1])$.
 
-When the "reset gate" is close to zero, the entire hidden state is thrown away and resetted with
+When the "reset gate" is close to zero, the entire hidden state is thrown away and reset with
 the current input gate state.
@@
 problem, large matrices multiplication is expensive. Word Embeddings provide an
-alternative procedure in which the words to be feeded to the network are encoded
+alternative procedure in which the words to be fed to the network are encoded
 in an $n$ dimensional space. The embedding layer can learn a word encoding such
 that words related to each other semanticaly, get similar encoded vectors after
-the transformation. Even some basic arithmetic is possible. For example, if you
-subtract the vector for "queen" from the vector for "king" you get the vector
+the transformation. Even some basic arithmetic is possible by adding and subtracting the euclidean distance of vectors. For example, if you
+subtract (the vector for "queen" - the vector from "woman") from the vector for "king" you get the vector
 for "man".
@@
 
-\includegraphics[width=0.7\columnwidth][width=\textwidth]{01}
+\begin{figure}[!ht]
+\includegraphics[width=\textwidth]{01}
+\centering
+\caption{Three embedding vector spaces showing how euclidean distance is following the semantic relations for: a) male and female. b) verb tenses. c) countries and capitals }
+\end{figure}
 
@@
 
-\subsection{Tied Weights}
-\label{sec:tiedweights}
-
 \section{Regularization}
@@
 
+\subsection{Tied Weights}
+\label{sec:tiedweights}
+
+Another regularization technique addresses the parameters in the encoder and the decoder.
+In our models, both of these layers have the vocabulary size as one of their dimensions.
+The encoder transforms the (vocabulary sized) one-hot encoding into a word encoding while the decoder transforms the hidden state into a softmax as big as the vocabulary size.
+These sizes can get huge as the vocabulary can contain from 20k to 800k with languages with many compound words (such as Finnish), which means that overfitting is quite likely as many of the parameters included in encoder and decoder may be redundant.
+
+In order to help solving this issue, Press O.
+et al \cite{press16} proposed sharing the same weights for encoder and decoder, as one is the same size as the transpose of the other and both perform a translation between a dense encoded layer and a sparse one.
+They showed in their work how this technique can reduce the number of parameters in the model without hurting the performance while lowering the perplexity score.
+
 \section{Initialization}
@@
 
+The starting values for the model parameters play a significant role in the training process for RNNs.
+The reason is mainly due to the depth these models reach while propagating through time.
+As the hidden state travels from time step to the next, an activation function is applied.
+The compound effect of performing the activation function several times tend to take the hidden state value towards the saturation zone of the activation functions.
+In order to avoid the saturation of the hidden units, the starting values must be chose carefully to ensure a correct behavior from the beginning, starting the training with as many active units as possible.
+
 \subsection{Xavier}
 \label{sec:xavier}
+
+Xavier initialization \cite{xavier10} was designed for the sigmoid and tanh functions. This is the initialization relevant to us since we have applied such functions more than the RELU.
+Xavier initialization consists on starting the hidden state values as a random distribution with variance as follows:
+$$var = \sqrt{\frac{1}{size_h}}$$
+
+Where $size_h$ is the number of hidden units.
+This ensures that the variance at the next hidden state is the same as the variance at the previous one.
+Maximizing the probability of having hidden units close to the mean and therefore out of the saturation zones.
 
@@
 
-\section{Performance by Model Architecture}
+\subsection{Performance by Model Architecture}
 \label{sec:perf_by_model_arch}
 
-\begin{figure}[ht]
+\begin{figure}[!ht]
 \includegraphics[width=0.7\columnwidth]{sr-perf-by-model-architecture}
@@
 
-\section{Performance by number of hidden layers}
+\subsection{Performance by number of hidden layers}
 \label{sec:perf_by_n_layers}
 
-\begin{figure}[ht]
+\begin{figure}[!ht]
 \includegraphics[width=0.7\columnwidth]{sr-perf-by-hidden-lstm}
@@
 
-\section{Performance by number of hidden units}
+\subsection{Performance by number of hidden units}
 \label{sec:perf_by_n_hidden}
 
-\begin{figure}[ht]
+\begin{figure}[!ht]
 \includegraphics[width=0.7\columnwidth]{sr-perf-by-hidden-units-lstm}
@@
 
-\section{Performance by Dropout}
+\subsection{Performance by Dropout}
 \label{sec:perf_by_dropout}
 
-\begin{figure}[ht]
+\begin{figure}[!ht]
 \includegraphics[width=0.7\columnwidth]{sr-perf-by-dropout-lstm}
@@
 
-\section{Performance by Embedding Dimensions}
+\subsection{Performance by Embedding Dimensions}
 \label{sec:perf_by_embedding}
 
-\begin{figure}[ht]
+\begin{figure}[!ht]
 \includegraphics[width=0.7\columnwidth]{sr-perf-by-embedding-lstm}
@@
 
-\section{Performance by Shortlist Dimension}
+\subsection{Performance by Shortlist Dimension}
 \label{sec:perf_by_shortlist}
 
-\begin{figure}[ht]
+\begin{figure}[!ht]
 \includegraphics[width=0.7\columnwidth]{sr-perf-by-shortlist-lstm}
@@
 
-\section{Tied Weights}
+\subsection{Performance by Tied Weights}
 \label{sec:perf_by_tiedweights}
 
-\begin{figure}[ht]
-\includegraphics[width=0.7\columnwidth]{sr-perf-by-model-tied-weights}
+We measured performance with tied weights, word embedding of size 200 and 2 hidden layers of size equal to the word embedding (200), this is a requirement for the encoder and decoder must have the same size in order to tie their weights.
+We trained to five epochs in batches of 200 sentences at a time.
+
+For the implementation of Tied Weights we were expecting lower perplexity levels as well as a small improvement or no change on the other metrics, however the LSTM and GRU models didn't show any perplexity improvements while the other metrics got worse.
+Our implementation of tied weights had a negative impact.
+For the tanh model as small perplexity improvement is shown but the other metrics are very unstable to draw any conclusions from them.
+
+\begin{figure}[!ht]
+\includegraphics[width=\textwidth, height=3cm]{lstm-tw}
+\centering
+\end{figure}
+\begin{figure}[!ht]
+\includegraphics[width=\textwidth, height=3cm]{gru-tw}
+\centering
+\end{figure}
+\begin{figure}[!ht]
+\includegraphics[width=\textwidth, height=3cm]{tanh-tw}
 \centering
@@
 
-\section{Xavier Initialization}
+\subsection{Performance by Xavier Initialization}
 \label{sec:perf_by_xavier_init}
 
-\begin{figure}[ht]
-\includegraphics[width=0.7\columnwidth]{sr-perf-by-model-xavier-init}
+We measured performance with Xavier initialization, word embedding of size 500 and 2 hidden layers of size 200.
+We trained to five epochs in batches of 200 sentences at a time.
+
+In all three models we get worse results when implementing Xavier initialization.
+The perplexity is not modified by much but METEOR, ROUGE and Naive Accuracy are clearly worse.
+Analysis of the activation function values across training steps would be needed to determine why Xavier initialization wasn't able to improve the results.
+
+\begin{figure}[!ht]
+\includegraphics[width=\textwidth, height=3cm]{lstm-xi}
 \centering
-\caption{Validation Set Performance by Model on WikiText for Xavier Initialization}
 \end{figure}
-
-\section{Kaiming Initialization}
-\label{sec:perf_by_kaiming_init}
-
-\begin{figure}[ht]
-\includegraphics[width=0.7\columnwidth]{sr-perf-by-model-kaiming-init}
+\begin{figure}[!ht]
+\includegraphics[width=\textwidth, height=3cm]{gru-xi}
+\centering
+\end{figure}
+\begin{figure}[!ht]
+\includegraphics[width=\textwidth, height=3cm]{tanh-xi}
 \centering
@@
 
+\bibitem{Bengio94}
+  http://ai.dinfo.unifi.it/paolo//ps/tnn-94-gradient.pdf
+
+\bibitem{hochreiter97}
+  http://www.bioinf.jku.at/publications/older/2604.pdf
+
+\bibitem{schwenk05}
+  ftp://tlp.limsi.fr/public/emnlp05.pdf. Accessed 6 Nov. 2018.
+
 \bibitem{Milkolov10}
@@
 
-\bibitem{Bengio94}
-  http://ai.dinfo.unifi.it/paolo//ps/tnn-94-gradient.pdf
+\bibitem{xavier10}
+  Understanding the difficulty of training deep feedforward neural networks - Glorot, Xavier \& Bengio Yoshua. 2010
+
+\bibitem{hinton2012}
+  https://arxiv.org/pdf/1207.0580.pdf
+
+\bibitem{cho2014}
+  https://arxiv.org/pdf/1406.1078v3.pdf
 
@@
 
-\bibitem{hochreiter97}
-  http://www.bioinf.jku.at/publications/older/2604.pdf
-
-\bibitem{cho2014}
-  https://arxiv.org/pdf/1406.1078v3.pdf
-
-\bibitem{schwenk05}
-  ftp://tlp.limsi.fr/public/emnlp05.pdf. Accessed 6 Nov. 2018.
-
-\bibitem{hinton2012}
-  https://arxiv.org/pdf/1207.0580.pdf
+\bibitem{press16}
+  Using the Output Embedding to Improve Language Models - Press, Ofir \& Wolf, Lior. 2016
 

### 2018-12-12 paper/paper.tex
paper: Reflow Xavier Initialization section
paper: Tidy up wording a little on Xavier Initialization
paper: Reflow section on tied weights
paper: Clarify missing word counts in gigaword/wikitext

@@
 
-This dataset is based on articles from Wikipedia, containing 1,756,924 tokens in
+This dataset is based on articles from Wikipedia, containing 1,205,450 tokens in
 the training set, 184,354 tokens in the validation set and 207,080 tokens
-in the test set. The combined dictionary for all three sets contains 28,755
+in the test set. The combined dictionary for all three sets contains 21,190
 unique tokens.
@@
 The second dataset we used was Gigaword. This dataset is based on XXX,
-containing 110,382,986 tokens in the training set, 549,0454 tokens in the
-validation set and XXX tokens in the test set. The combined dictionary
-contains XXX unique tokens.
+containing 8,658,023 tokens in the training set, and 341,317 tokens in the
+validation set. The combined dictionary contains 59760 unique tokens.
 
@@
 
-XXX
+"five time world champion kwan withdrew from the US figure skating championships
+on Wednesday, but will petition US skating officials for the chance to compete
+\emph{at the Turin Olympics}."
+
+"US business leaders lashed out wednesday at legislation that would penalize companies
+\epmh{for employing illegal immigrants}."
 
@@
 
-We measured performance with tied weights, word embedding of size 200 and 2 hidden layers of size equal to the word embedding (200), this is a requirement for the encoder and decoder must have the same size in order to tie their weights.
-We trained to five epochs in batches of 200 sentences at a time.
+We measured performance with tied weights, word embedding of size 200 and 2
+hidden layers of size equal to the word embedding (200), this is a requirement
+for the encoder and decoder must have the same size in order to tie their
+weights. We trained to five epochs in batches of 200 sentences at a time.
 
-For the implementation of Tied Weights we were expecting lower perplexity levels as well as a small improvement or no change on the other metrics, however the LSTM and GRU models didn't show any perplexity improvements while the other metrics got worse.
-Our implementation of tied weights had a negative impact.
-For the tanh model as small perplexity improvement is shown but the other metrics are very unstable to draw any conclusions from them.
+For the implementation of Tied Weights we were expecting lower perplexity levels
+as well as a small improvement or no change on the other metrics, however the
+LSTM and GRU models didn't show any perplexity improvements while the other
+metrics got worse. Our implementation of tied weights had a negative impact. For
+the tanh model as small perplexity improvement is shown but the other metrics
+are very unstable to draw any conclusions from them.
 
@@
 
-In all three models we get worse results when implementing Xavier initialization.
-The perplexity is not modified by much but METEOR, ROUGE and Naive Accuracy are clearly worse.
-Analysis of the activation function values across training steps would be needed to determine why Xavier initialization wasn't able to improve the results.
+In all three models we get worse results when implementing Xavier initialization. Xavier Initialization does not
+affect perplexity that much, but but METEOR, ROUGE and Naive Accuracy performed much worse than the baseline.
+Analysis of the activation function values across training steps would be needed to determine why Xavier
+initialization wasn't able to improve the results.
 

### 2018-12-12 paper/paper.tex
paper: Add comment on edit distance
paper: Reflow line
paper: Reflow tied weights and xavier init

@@
 words and $S, D, I$ are substitutions, deletions, insertions.
+This is similar to the Levenstein distance between documents.
 
@@
 paper consists of a single linear layer and non-linear
-activation (in our case, the $tanh$ function, $ \frac{e^{2x} - 1}{e^{2x} + 1} $, with domain $[ -\infty, \infty ]$ and
+activation (in our case, the $tanh$ function,
+$ \frac{e^{2x} - 1}{e^{2x} + 1} $, with domain $[ -\infty, \infty ]$ and
 range $ [-1, 1] $. The inputs to the network are
@@
 
-Another regularization technique addresses the parameters in the encoder and the decoder.
-In our models, both of these layers have the vocabulary size as one of their dimensions.
-The encoder transforms the (vocabulary sized) one-hot encoding into a word encoding while the decoder transforms the hidden state into a softmax as big as the vocabulary size.
-These sizes can get huge as the vocabulary can contain from 20k to 800k with languages with many compound words (such as Finnish), which means that overfitting is quite likely as many of the parameters included in encoder and decoder may be redundant.
+Another regularization technique addresses the parameters in the encoder and the
+decoder. In our models, both of these layers have the vocabulary size as one of
+their dimensions. The encoder transforms the (vocabulary sized) one-hot encoding
+into a word encoding while the decoder transforms the hidden state into a
+softmax as big as the vocabulary size. These sizes can get huge as the
+vocabulary can contain from 20k to 800k with languages with many compound words
+(such as Finnish), which means that overfitting is quite likely as many of the
+parameters included in encoder and decoder may be redundant.
 
-In order to help solving this issue, Press O.
-et al \cite{press16} proposed sharing the same weights for encoder and decoder, as one is the same size as the transpose of the other and both perform a translation between a dense encoded layer and a sparse one.
-They showed in their work how this technique can reduce the number of parameters in the model without hurting the performance while lowering the perplexity score.
+In order to help solving this issue, Press O. et al \cite{press16} proposed
+sharing the same weights for encoder and decoder, as one is the same size as the
+transpose of the other and both perform a translation between a dense encoded
+layer and a sparse one. They showed in their work how this technique can reduce
+the number of parameters in the model without hurting the performance while
+lowering the perplexity score.
 
@@
 
-The starting values for the model parameters play a significant role in the training process for RNNs.
-The reason is mainly due to the depth these models reach while propagating through time.
-As the hidden state travels from time step to the next, an activation function is applied.
-The compound effect of performing the activation function several times tend to take the hidden state value towards the saturation zone of the activation functions.
-In order to avoid the saturation of the hidden units, the starting values must be chose carefully to ensure a correct behavior from the beginning, starting the training with as many active units as possible.
+The starting values for the model parameters play a significant role in the
+training process for RNNs. The reason is mainly due to the depth these models
+reach while propagating through time. As the hidden state travels from time step
+to the next, an activation function is applied. The compound effect of
+performing the activation function several times tend to take the hidden state
+value towards the saturation zone of the activation functions. In order to avoid
+the saturation of the hidden units, the starting values must be chose carefully
+to ensure a correct behavior from the beginning, starting the training with as
+many active units as possible.
 
@@
 
-Xavier initialization \cite{xavier10} was designed for the sigmoid and tanh functions. This is the initialization relevant to us since we have applied such functions more than the RELU.
-Xavier initialization consists on starting the hidden state values as a random distribution with variance as follows:
-$$var = \sqrt{\frac{1}{size_h}}$$
+Xavier initialization \cite{xavier10} was designed for the sigmoid and tanh
+functions. This is the initialization relevant to us since we have applied such
+functions more than the RELU. Xavier initialization consists on starting the
+hidden state values as a random distribution with variance as follows: $$var =
+\sqrt{\frac{1}{size_h}}$$
 
-Where $size_h$ is the number of hidden units.
-This ensures that the variance at the next hidden state is the same as the variance at the previous one.
-Maximizing the probability of having hidden units close to the mean and therefore out of the saturation zones.
+Where $size_h$ is the number of hidden units. This ensures that the variance at
+the next hidden state is the same as the variance at the previous one.
+Maximizing the probability of having hidden units close to the mean and
+therefore out of the saturation zones.
 

### 2018-12-12 paper/paper.tex
paper: Remove extra space
paper: Add conclusions, future work section.

@@
 
+Experimentation with model hyperparameters that increase or decrease information
+retention (such as number of hidden layers or their dimensionality) peformed in line
+with expectations. In general, increasing the amount of information stored by the network
+allowed it to generate more marginally more realistic language with a lower degree of
+perplexity at the output layers.
+
+Usage of trainable word embeddings at the encoding layers outperformed our expectations -
+we found that there were no situations where using word embeddings did not result in both
+a computational performance and natural language generation performance improvement. In fact,
+word embeddings resulted in the most significant perforamnce improvement out of any of the
+modifications that we made to the baseline models. We also found that using output shortlists
+of length 5000 or above for the WikiText models resulted in scores competitve with the baseline
+on both METEOR and ROUGE, but with twice to fource times less computational complexity because
+not as many dot products need to be performed at the output layer for each word prediction. However,
+if one had to choose between shortlisting and trainable word embeddings at the input layer, embeddings
+are able to reduce compute complexity much more whilst also improving accuracy, something that
+did not happen with shortlists at the output layer.
+
+We found that like the paper suggested, usage of regularization techniques did not assist in improving
+either perplexity or scores on both METEOR and ROUGE. This was even true in the case of Dropout,
+which was proposed after the 2010 paper by Milkolov was published. Such a result was surprising to us
+and perhaps requires further investigation and testing. For instance, Zaremba & Vinyals found that
+applying dropout on non-recurrent connections could be an effective regularizer \cite{zaremba2015},
+though this is different to the way in which we implemented dropout, which was on both the recurrent
+and non-recurrent connections. This may have impacted the network's ability to remember crucial
+information during training that it would allow it to make better predictions and optimize better.
+
+Finally, we found that using Tied Weights and Xavier Initialization did not assist on any of our
+natural language generation metrics and in fact made performance worse. It was surprising to see
+that Xavier Initialization made the model perform more poorly, so perhaps further analysis is required
+as to the state of the activations between the recurrent connections to determine if they are more
+or less saturated during training.
+
 \section{Future Work}
@@
 
-\begin{thebibliography}{9}
-\bibitem{nano3}
-  K. Grove-Rasmussen og Jesper Nygård,
-  \emph{Kvantefænomener i Nanosystemer}.
-  Niels Bohr Institute \& Nano-Science Center, Københavns Universitet
+In this paper, we did not have the opportunity to evaluate some of the methods identified in our
+original literature review. For instance, we were not able to implement a working version of
+Batch Normalization \cite{ioffe} which has been shown to be effective in the context of LSTMs \cite{courville2017}.
 
-\bibitem{Bengio94}
-  http://ai.dinfo.unifi.it/paolo//ps/tnn-94-gradient.pdf
+We also did not have the opportunity to examine Adaptive Importance Sampling \cite{bengio08}. Doing so
+likely would have taken a substantial amount of time to implement as it stands in as an alternative
+to full backpropagation. However, it would be useful to see if the technique is effective in
+reducing the training time of the RNN models by merely approximating the converged state.
 
-\bibitem{hochreiter97}
-  http://www.bioinf.jku.at/publications/older/2604.pdf
-
-\bibitem{schwenk05}
-  ftp://tlp.limsi.fr/public/emnlp05.pdf. Accessed 6 Nov. 2018.
-
-\bibitem{Milkolov10}
-  Milkolov T., et al, Recurrent neural network based language model, INTERSPEECH 2010
-
-\bibitem{xavier10}
-  Understanding the difficulty of training deep feedforward neural networks - Glorot, Xavier \& Bengio Yoshua. 2010
-
-\bibitem{hinton2012}
-  https://arxiv.org/pdf/1207.0580.pdf
-
-\bibitem{cho2014}
-  https://arxiv.org/pdf/1406.1078v3.pdf
-
-\bibitem{colah2015}
-  http://colah.github.io/posts/2015-08-Understanding-LSTMs/
-
-\bibitem{press16}
-  Using the Output Embedding to Improve Language Models - Press, Ofir \& Wolf, Lior. 2016
-
-\end{thebibliography}
+Finally, we did not get the opportunity to implement the open vocabulary approach, which would allow
+the network to fall back to predicting character data if an UNK token was predicted \cite{ling2016}. Such
+an approach again would have required substantial modification to the baseline implementation that
+we used, since the baseline predicts one-hot encoded words in batches and is not set up to deal with
+character encodings.
 

### 2018-12-12 paper/paper.tex
paper: Make wikitext and gigaword subsections
paper: Fix emph
paper: Fix duplicate label
paper: Escape &

@@
 
-\section{WikiText}
+\subsection{WikiText}
 \label{sec:wikitext}
@@
 
-\section{Gigaword}
+\subsection{Gigaword}
 \label{sec:gigaword}
@@
 "US business leaders lashed out wednesday at legislation that would penalize companies
-\epmh{for employing illegal immigrants}."
+\emph{for employing illegal immigrants}."
 
@@
 \subsection{RNN}
-\label{sec:lstm}
+\label{sec:rnn}
 
@@
 which was proposed after the 2010 paper by Milkolov was published. Such a result was surprising to us
-and perhaps requires further investigation and testing. For instance, Zaremba & Vinyals found that
+and perhaps requires further investigation and testing. For instance, Zaremba 'I\&' Vinyals found that
 applying dropout on non-recurrent connections could be an effective regularizer \cite{zaremba2015},

### 2018-12-12 paper/paper.tex
paper: Add external bibliography
paper: Improve citations

@@
 
-One of the first papers applying RNNs to Language Modeling was \cite{Milkolov10}. But since 2010, new RNN approaches have shown
-improvement on the state-of-the-art. We examine these new architectures and
-techniques for future work discussed in the Milkolov paper and analyze whether their
-results indicate that improvements can be made on the baseline set by the paper.
+One of the first papers applying RNNs to Language Modeling was by Milkolov and others \cite{Milkolov10}.
+But since 2010, new RNN approaches have shown improvement on the state-of-the-art.
+We examine these new architectures and techniques for future work discussed in the
+Milkolov paper and analyze whether their results indicate that improvements can be made
+on the baseline set by the paper.
 
@@
 Unfortunately, Elman Network RNNs are not able to retain all the optimal
-contextual in practice. The reasons for this are given by \cite{Bengio94},
+contextual in practice. The reasons for this are given by Bengio\cite{Bengio94},
 but in summary the issue is that as we backpropagate through time the gradients
@@
 
+\bibliographystyle{plain}
+\bibliography{bibliography.bib}
 \end{document}

### 2018-12-12 paper/paper.tex
paper: Add citations for nlg-eval and pytorch lm
paper: Make the images slightly bigger

@@
 \begin{figure}[!ht]
-\includegraphics[width=0.7\columnwidth]{sr-eda-wikitext-train-words}
-\includegraphics[width=0.7\columnwidth]{sr-eda-wikitext-valid-words}
+\includegraphics[width=0.8\columnwidth]{sr-eda-wikitext-train-words}
+\includegraphics[width=0.8\columnwidth]{sr-eda-wikitext-valid-words}
 \centering
@@
 \begin{figure}[!ht]
-\includegraphics[width=0.7\columnwidth]{sr-eda-wikitext-bigrams}
-\includegraphics[width=0.7\columnwidth]{sr-eda-wikitext-trigrams}
+\includegraphics[width=0.8\columnwidth]{sr-eda-wikitext-bigrams}
+\includegraphics[width=0.8\columnwidth]{sr-eda-wikitext-trigrams}
 \centering
@@
 \begin{figure}[!ht]
-\includegraphics[width=0.7\columnwidth]{sr-eda-gigaword-train-words}
-\includegraphics[width=0.7\columnwidth]{sr-eda-gigaword-valid-words}
+\includegraphics[width=0.8\columnwidth]{sr-eda-gigaword-train-words}
+\includegraphics[width=0.8\columnwidth]{sr-eda-gigaword-valid-words}
 \centering
@@
 \begin{figure}[!ht]
-\includegraphics[width=0.7\columnwidth]{sr-eda-gigaword-bigrams}
-\includegraphics[width=0.7\columnwidth]{sr-eda-gigaword-trigrams}
+\includegraphics[width=0.8\columnwidth]{sr-eda-gigaword-bigrams}
+\includegraphics[width=0.8\columnwidth]{sr-eda-gigaword-trigrams}
 \centering
@@
 
+The actual models were implemented as an extension to the PyTorch
+WikiText RNN Language Modelling Example \cite{pytorchlm}. Metrics were
+evaluated using the NLG-Eval toolkit by Sharma and others \cite{nlgeval}.
+
 \subsection{Performance by Model Architecture}
@@
 \begin{figure}[!ht]
-\includegraphics[width=0.7\columnwidth]{sr-perf-by-model-architecture}
+\includegraphics[width=0.8\columnwidth]{sr-perf-by-model-architecture}
 \centering
@@
 \begin{figure}[!ht]
-\includegraphics[width=0.7\columnwidth]{sr-perf-by-hidden-lstm}
-\includegraphics[width=0.7\columnwidth]{sr-perf-by-hidden-gru}
-\includegraphics[width=0.7\columnwidth]{sr-perf-by-hidden-rnn}
+\includegraphics[width=0.8\columnwidth]{sr-perf-by-hidden-lstm}
+\includegraphics[width=0.8\columnwidth]{sr-perf-by-hidden-gru}
+\includegraphics[width=0.8\columnwidth]{sr-perf-by-hidden-rnn}
 \centering
@@
 \begin{figure}[!ht]
-\includegraphics[width=0.7\columnwidth]{sr-perf-by-hidden-units-lstm}
-\includegraphics[width=0.7\columnwidth]{sr-perf-by-hidden-units-gru}
-\includegraphics[width=0.7\columnwidth]{sr-perf-by-hidden-units-rnn}
+\includegraphics[width=0.8\columnwidth]{sr-perf-by-hidden-units-lstm}
+\includegraphics[width=0.8\columnwidth]{sr-perf-by-hidden-units-gru}
+\includegraphics[width=0.8\columnwidth]{sr-perf-by-hidden-units-rnn}
 \centering
@@
 \begin{figure}[!ht]
-\includegraphics[width=0.7\columnwidth]{sr-perf-by-dropout-lstm}
-\includegraphics[width=0.7\columnwidth]{sr-perf-by-dropout-gru}
-\includegraphics[width=0.7\columnwidth]{sr-perf-by-dropout-rnn}
+\includegraphics[width=0.8\columnwidth]{sr-perf-by-dropout-lstm}
+\includegraphics[width=0.8\columnwidth]{sr-perf-by-dropout-gru}
+\includegraphics[width=0.8\columnwidth]{sr-perf-by-dropout-rnn}
 \centering
@@
 \begin{figure}[!ht]
-\includegraphics[width=0.7\columnwidth]{sr-perf-by-embedding-lstm}
-\includegraphics[width=0.7\columnwidth]{sr-perf-by-embedding-gru}
-\includegraphics[width=0.7\columnwidth]{sr-perf-by-embedding-rnn}
+\includegraphics[width=0.8\columnwidth]{sr-perf-by-embedding-lstm}
+\includegraphics[width=0.8\columnwidth]{sr-perf-by-embedding-gru}
+\includegraphics[width=0.8\columnwidth]{sr-perf-by-embedding-rnn}
 \centering
@@
 \begin{figure}[!ht]
-\includegraphics[width=0.7\columnwidth]{sr-perf-by-shortlist-lstm}
-\includegraphics[width=0.7\columnwidth]{sr-perf-by-shortlist-gru}
-\includegraphics[width=0.7\columnwidth]{sr-perf-by-shortlist-rnn}
+\includegraphics[width=0.8\columnwidth]{sr-perf-by-shortlist-lstm}
+\includegraphics[width=0.8\columnwidth]{sr-perf-by-shortlist-gru}
+\includegraphics[width=0.8\columnwidth]{sr-perf-by-shortlist-rnn}
 \centering

### 2018-12-12 paper/paper.tex
paper: Fix some typos

@@
 with this model is that the context window is of a fixed size, leading
-to a horizon problem. The Horizon problem occur because a if there were a word
+to a horizon problem. The horizon problem occurs because a if there were a word
 in the word sequence which is extremely predictive of the $(n + 1)$th
-word away from the current word, that would would simply not be taken
+word away from the current word, that word would simply not be taken
 into account in that model. Worse still, the fixed size context window
@@
 
-The second dataset we used was Gigaword. This dataset is based on XXX,
+The second dataset we used was Gigaword. This dataset is based on news articles,
 containing 8,658,023 tokens in the training set, and 341,317 tokens in the
@@
 
-Samples from the Dataset include:
+Samples from the Dataset include (expected predictions highlighted):
 
-"five time world champion kwan withdrew from the US figure skating championships
+"five time world champion Kwan withdrew from the US figure skating championships
 on Wednesday, but will petition US skating officials for the chance to compete
@@
 
-"US business leaders lashed out wednesday at legislation that would penalize companies
+"US business leaders lashed out Wednesday at legislation that would penalize companies
 \emph{for employing illegal immigrants}."
@@
 words and $S, D, I$ are substitutions, deletions, insertions.
-This is similar to the Levenstein distance between documents.
+This is similar to the Levensteihn distance between documents.
 

### 2018-12-13 paper/paper.tex
paper: Throw-away line about why we care about the data
review of paper and fixing the bibliography

@@
 with this model is that the context window is of a fixed size, leading
-to a horizon problem. The horizon problem occurs because a if there were a word
+to a horizon problem. The horizon problem occurs because if there were a word
 in the word sequence which is extremely predictive of the $(n + 1)$th
@@
 
-In contrast, RNN's seek to solve this problem by introducing a Neural Network
+In contrast, RNNs seek to solve this problem by introducing a Neural Network
 architecture with recurrent connections, meaning that as the RNN makes predictions
@@
 
+We build on top of the paper by performing experiments with new model types
+and improvements and measure their effectiveness against certain well-known datasets.
+
 \section{Data}
@@
 The first dataset we use is the WikiText dataset. This is a smaller dataset than
-Gigaword and WSJ'92, but allowed us to
+Gigaword and WSJ, but allowed us to
 iterate faster and run more experiments.
@@
 words and $S, D, I$ are substitutions, deletions, insertions.
-This is similar to the Levensteihn distance between documents.
+This is similar to the Levenshtein Distance between documents (Haldar et al. \cite{haldar11}).
 
-However, such a metric doesn't take int account the fact that
+However, such a metric doesn't take into account the fact that
 a recognized sequence may differ in length to a reference
@@
 Improving on this situation, ROUGE measures based on the
-longest matching subsequences, overlapping pairs and n-gram
+longest matching sub-sequences, overlapping pairs and n-gram
 co-occurrences.
@@
 Precision over uni-grams, the weight for recall being higher than Precision,
-meaning that false negatives (having a low probability for the true word trying
+meaning that false negatives (having a low probability for the true word we are  trying
 to predict) will have a big impact on the metric (the score will be lower as the
@@
 
-More formally, Precision over an uni-gram is computed as: $$P={\frac
+Formally, Precision over an uni-gram is computed as: $$P={\frac
 {m}{w_{{t}}}}$$ Where m is the number of uni-grams in the candidate translation
@@
 but in summary the issue is that as we backpropagate through time the gradients
-from states "further back" in the history start to get vanishingly small, meaning
+from states "further back" in the history start to get smaller and smaller, meaning
 that they can be easily disrupted by noise or newer information. This means
@@
 The LSTM cell uses four "gates" to manage the state of the memory. Each of these
-gates comprises a set of weights and activations, descibed below \cite{colah2015}.
+gates comprises a set of weights and activations, described below \cite{colah2015}.
 
@@
 The last gate is an "output" gate which decides which parts of the hidden state are
-taken on account to compute the output. The hidden state is first activated by sigmoid to determin
+taken on account to compute the output. The hidden state is first activated by sigmoid to determine
 the mask ($o_t$), then the hidden layer is activated by tanh to push the values between $[-1, 1]$, then
-finally multipled by the mask $o_t$ in order to determine the output of the network.
+finally multiplied by the mask $o_t$ in order to determine the output of the network.
 
@@
 by virtue of being combined through a single linear layer and activation layer with the input
-state on every timestep. Instead, the hidden state is more "explicitly" modified when the
+state on every time step. Instead, the hidden state is more "explicitly" modified when the
 optimized input and forget gates explicitly indicate that information should be added or removed
@@
 
-Note that the hidden state accuracy benefits here come with a computational compleixty cost - every
+Note that the hidden state accuracy benefits here come with a computational complexity cost - every
 LSTM cell has three internal linear layers and five activations, which we pay for both in the forward
@@
 
-Traditionally, the Language Models were feeded each word as a One-Hot encoded vector as large as the
+Traditionally, the Language Models were fed each word as a One-Hot encoded vector as large as the
 vocabulary is. Since vocabularies can be really large (20k to 800k for languages
 with many compounded words such as Finnish), this presents a computational
-problem, large matrices multiplication is expensive. Word Embeddings provide an
-alternative procedure in which the words to be fed to the network are encoded
+problem, matrix multiplication can get very  expensive quite fast as the size of the matrices being multiplied grow. Word Embeddings provide an
+alternative procedure to One-Hot in which the words to be fed to the network are encoded
 in an $n$ dimensional space. The embedding layer can learn a word encoding such
-that words related to each other semanticaly, get similar encoded vectors after
+that words semantically related to each other, get similar encoded vectors after
 the transformation. Even some basic arithmetic is possible by adding and subtracting the euclidean distance of vectors. For example, if you
-subtract (the vector for "queen" - the vector from "woman") from the vector for "king" you get the vector
+subtract (the vector for "queen" minus the vector from "woman") from the vector for "king" you get the vector
 for "man".
@@
 the RNN, LSTM or GRU cell phase both at the inference and training time. The
-comptutation of the embedding layer can be heavily optimized because the relevant
+computation of the embedding layer can be heavily optimized because the relevant
 input will only have a non-zero value in a single vector component.
@@
 The overall theory behind shortlists is that any optimal model operates
-on a pareto principle, 20% of the words which are the most frequent will
+on a Pareto principle, 20% of the words which are the most frequent will
 be predicted roughly 80% of the time. Since the computation of the output
@@
 cycle, the user of the model is paying for the capacity to predict words which
-will seldomly be predicted. A good shortlist should be able to substantially
+will seldom be predicted. A good shortlist should be able to substantially
 reduce the computational complexity of the model whilst still retaining a
@@
 improved by using "standard regularization procedures" (we assume that
-the paper must have been referring to L1 or L2 regualrization based on the
+the paper must have been referring to L1 or L2 regularization based on the
 techniques in existence at the time) \cite{Milkolov10}. We propose that a
-popular regularization technique known as "Dropout" may be of some assitance.
+popular regularization technique known as "Dropout" may be of some assistance.
 
@@
 their dimensions. The encoder transforms the (vocabulary sized) one-hot encoding
-into a word encoding while the decoder transforms the hidden state into a
+into a word embedding while the decoder transforms the hidden state into a
 softmax as big as the vocabulary size. These sizes can get huge as the
-vocabulary can contain from 20k to 800k with languages with many compound words
-(such as Finnish), which means that overfitting is quite likely as many of the
+vocabulary can contain from 20k to 800k, which means that overfitting is quite likely as many of the
 parameters included in encoder and decoder may be redundant.
 
-In order to help solving this issue, Press O. et al \cite{press16} proposed
+In order to help solving this issue, Press O. et al. \cite{press16} proposed
 sharing the same weights for encoder and decoder, as one is the same size as the
@@
 value towards the saturation zone of the activation functions. In order to avoid
-the saturation of the hidden units, the starting values must be chose carefully
-to ensure a correct behavior from the beginning, starting the training with as
+the saturation of the hidden units, the starting values must be chose carefully. The goal is then 
+to ensure a correct behavior from the beginning, starting the training process with as
 many active units as possible.
@@
 functions. This is the initialization relevant to us since we have applied such
-functions more than the RELU. Xavier initialization consists on starting the
+functions more than the RELU alternative. Xavier initialization consists on starting the
 hidden state values as a random distribution with variance as follows: $$var =
@@
 We measured differences in performance with 60, 90, 150 and 200 hidden units per
-RNN cell for each model architecture with 1 RNN cell, trained
+RNN cell for each model architecture with 1 RNN layer, trained
 to five epochs in batches of 200 sentences at a time.
@@
 
-On both LSTM and GRU models, any use of encoder embeddings noticably outperformed
+On both LSTM and GRU models, any use of encoder embeddings noticeably outperformed
 one-hot encoding on almost all metrics, except METEOR for LSTMs. It seems to be
@@
 metrics got worse. Our implementation of tied weights had a negative impact. For
-the tanh model as small perplexity improvement is shown but the other metrics
-are very unstable to draw any conclusions from them.
+the tanh model a small perplexity improvement is shown but the other metrics
+are too unstable to draw any conclusions from them.
 
@@
 Experimentation with model hyperparameters that increase or decrease information
-retention (such as number of hidden layers or their dimensionality) peformed in line
+retention (such as number of hidden layers or their dimensionality) performed in line
 with expectations. In general, increasing the amount of information stored by the network
-allowed it to generate more marginally more realistic language with a lower degree of
+allowed it to generate (marginally) more realistic language with a lower degree of
 perplexity at the output layers.
 
-Usage of trainable word embeddings at the encoding layers outperformed our expectations -
-we found that there were no situations where using word embeddings did not result in both
+Usage of trainable word embeddings at the encoding layers outperformed our expectations. We found that there were no situations where using word embeddings did not result in both
 a computational performance and natural language generation performance improvement. In fact,
-word embeddings resulted in the most significant perforamnce improvement out of any of the
+word embeddings resulted in the most significant performance improvement out of any of the
 modifications that we made to the baseline models. We also found that using output shortlists
 of length 5000 or above for the WikiText models resulted in scores competitve with the baseline
-on both METEOR and ROUGE, but with twice to fource times less computational complexity because
+on both METEOR and ROUGE, but with twice to four times less computational complexity because
 not as many dot products need to be performed at the output layer for each word prediction. However,
@@
 
+Additionally, it would be interesting to apply additional model architectures such as Bidirectional RNNs and Attention-based models.
+In particular, the latter have shown great performance improvements during the last years, achieving for example a new state-of-the-art result in machine translation in 2015 (Luong et al.
+\cite{luong15}) and also matching state-of-the-art results in Language Modeling with more than 4 times less parameters in comparison with other models (Salton et al.
+\cite{salton17}).
+
 Finally, we did not get the opportunity to implement the open vocabulary approach, which would allow
@@
 \bibliographystyle{plain}
-\bibliography{bibliography.bib}
+\bibliography{bibliography}
 \end{document}
